Wrinkling of a confined porous layer

### 2014-11-14 fig_plis/article_plis.tex
permeability : Figure : switch tube 1 and 2 in the sketch, use gray instead of yellow, add a second graph for glycerol 50%, use figure* Supplementary method

>>>>

\subsection*{Permeability measurements}
\begin{figure*}
	\includegraphics{permeability}
	\caption{Permeability measurements. (a) Schematic representation of the experiment. (b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2. Black line is the best exponential fit $Ae^{-t/\tau}$. (b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$. (c) Idem in 50\%w glycerol, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}
	\label{fig:permeability}
\end{figure*}

To measure the permeability of our gels, we adapt a method from \cite{VanDijk1986}. First, the gel of interest is formed at the lower extremity of a thin glass tube. The tube is held vertical in a water saturated atmosphere to prevent evaporation until the pH has reached the isoelectric point ($\approx\SI{1}{\hour}$ for our most common composition). Then, we dip the tube in a bath of acetic acid/acetate buffer at pH=4. The buffer has the same glycerol content as the gel solvent, so that the viscosity is constant throughout the experiment. Alongside the first tube (see Fig.~\ref{fig:permeability} a), we dip a second identical tube with no gel in it. A webcam (Logitech Webcam Pro 9000) records the rise of the liquid in each tube. Since the liquid height in tube 2 reaches its final value quickly, we measure directly $z_\infty-z(t)$.

Darcy law through the gel of permeability $\alpha$ and height $H$ writes
\begin{equation}
\frac{dz}{dt} = \frac{\alpha}{\eta}\frac{\Delta P}{H}
\end{equation}
where $\Delta P$ is the the hydrostatic pressure relative to $z_\infty$. We can thus write
\begin{equation}
\frac{d(z_\infty-z)}{dt} = -\frac{z_\infty-z}{\tau}\text{, with\,} \tau = \frac{\eta H}{\alpha \rho g}.
\end{equation}

By fitting $z_\infty-z(t)$ by an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$. Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosities and low permeability. Obtaining $z_\infty$ without tube 2 would be time prohibitive.
